Effects of Reynolds Number on the Energy Conversion and Near-Wake Dynamics of a High Solidity Vertical-Axis Cross-Flow Turbine

### 2016-01-21 paper.tex
Do not put dash before unit abbreviations

Blackwell \emph{et al.} \cite{Blackwell1976} investigated the effects of
Reynolds number on the performance of a 2-m diameter Darrieus vertical-axis
cross-flow wind turbine with NACA 0012 blades. By varying wind tunnel speed, the
turbine was made to operate at an approximately constant blade chord Reynolds
number $Re_c$ ranging from $1 \times 10^5$ to $3 \times 10^5$. In this regime,
the turbine power coefficient $C_P$ was shown to be sensitive to $Re_c$, with
sensitivity diminishing at the higher Reynolds numbers, especially for turbines
of lower solidity ($Nc/R$, where $N$ is the number of blades and $R$ is the
turbine's maximum radius). More recently, Polagye and Cavagnaro
\cite{Polagye2013b} observed significant Reynolds number sensitivity for a high
solidity helical cross-flow turbine, and Bravo \emph{et al.} \cite{Bravo2007}
observed the power coefficient of a straight-bladed VAWT to become Reynolds
number independent at $Re_c \approx 4 \times 10^5$. Bachant and Wosnik
\cite{Bachant2014} observed Reynolds number independence of the power
coefficient at the optimal tip speed ratio for a high solidity cross-flow
turbine at $Re_D \approx 10^6$ or $Re_{c,\mathrm{avg}} \approx 2 \times 10^5$.

@@

Scale effects on axial-flow or horizontal-axis wind turbines are better
understood, and investigators have methods for compensating. Krogstad and
Adaramola \cite{Krogstad2012a} observed Reynolds number independence of the
performance of a 0.9-m diameter axial-flow turbine rotor at $Re_D \approx 5
\times 10^5$ in wind tunnel tests. Their turbine blades had an S826 profile
along their entire span, a profile chosen for its $Re$-independence. Walker \emph{et
al.} \cite{Walker2014} similarly chose a NACA 63-618 foil for their axial-flow
turbine tests in a towing tank, which were performed at a mid-span Reynolds
number $Re_c = 4 \times 10^5$. McTavish \emph{et al.} \cite{McTavish2013} showed how
the near-wake expansion for an axial-flow rotor is increased at higher Reynolds
numbers, concluding that physical models should be designed with
$Re$-independence in mind if they are to be run at low $Re$. It is uncertain,
however, if these methods work equally well for cross-flow turbines.

@@

\caption{\hl {University of New Hampshire (UNH)}-Reference Vertical Axis Turbine (RVAT) turbine model. Turbine blades and struts made from NACA 0020
 profiles with 0.140-m chord. Note that the upper and lower mounting flanges have
 been excluded, as these were included in the tare drag measurements.}

@@

The turbine was mounted in a frame constructed from NACA 0020 sections, shown in
Figure~\ref{fig:exp-setup}. The turbine shaft ran up through the water surface,
coupled to a Kollmorgen \hl {AKM}
%please define AKM and RPM
  permanent magnet servo \hl {motor (Manufacturer, City, Country)} with a 20:1 gearbox,
providing precise control over shaft angular velocity. This servo was controlled by the tow
tank's main motion controller for high synchronization with the carriage motion,
thereby giving precise measurement and control of the tip speed ratio. An Interface
T8 200-Nm capacity rotary torque \hl {transducer (Manufacturer, City, Country)} was installed inline between the
servo and the turbine, and the servo was also mounted on a slewing ring bearing,
which allowed a redundant measurement of torque via an arm and load cell used to
counteract the turbine moment. The frame was mounted to the carriage via linear
guides, such that the total streamwise drag force was transferred to a pair of
Sentran ZB3 500 pound-force
 capacity S-beam \hl {load cells (Manufacturer, City, Country)}, providing the rotor drag
measurements, after a separately measured tare drag was subtracted in post-processing. Similarly, a tare torque was measured by rotating the turbine shaft
in air. Turbine angular and tow carriage linear position were measured using
quadrature encoder signals, with $10^5$ counts-per-rev for the turbine and 10 ${\mu}$m resolution for the carriage position. These~signals, along with
the torque and drag signals, were sampled at 2 kHz.

Wake velocity was measured using a Nortek Vectrino+ acoustic Doppler \hl {velocimeter
(ADV)  (Manufacturer, City, Country)}, which has an approximately 6-mm diameter sampling volume and sampled at
200 Hz. The probe was mounted on an automated positioning system, also
controlled by the tow tank's main motion controller. The ADV and \hl {DAQ (please define)} systems'
sampling times were synchronized by triggering the start of data acquisition via
a pulse sent from the motion controller. Additional details of the turbine and
experimental setup are described in \cite{Bachant2015-JoT}.

@@

Approximately 1500 tows were conducted for the study reported here; each tow was
used for a single data point on either a performance curve or wake map. Each
performance curve consisted of 31~tows, where during each tow, the mean turbine
tip speed ratio was held constant, ranging from 0.1 to 3.1 in 0.1 increments. Full
performance curve data were acquired for tow speeds from 0.4 to 1.2 m/s in 0.2-m/s increments, for which the turbine diameter and approximate blade chord Reynolds
number are presented in Table~\ref{tab:Re}. Performance was also measured for
$\lambda=1.9$ at tow speeds $[0.3, 0.5, 0.7, 0.9, 1.1, 1.3]$ m/s for two tows
each.

>>>>

Blackwell \emph{et al.} \cite{Blackwell1976} investigated the effects of
Reynolds number on the performance of a 2 m diameter Darrieus vertical-axis
cross-flow wind turbine with NACA 0012 blades. By varying wind tunnel speed, the
turbine was made to operate at an approximately constant blade chord Reynolds
number $Re_c$ ranging from $1 \times 10^5$ to $3 \times 10^5$. In this regime,
the turbine power coefficient $C_P$ was shown to be sensitive to $Re_c$, with
sensitivity diminishing at the higher Reynolds numbers, especially for turbines
of lower solidity ($Nc/R$, where $N$ is the number of blades and $R$ is the
turbine's maximum radius). More recently, Polagye and Cavagnaro
\cite{Polagye2013b} observed significant Reynolds number sensitivity for a high
solidity helical cross-flow turbine, and Bravo \emph{et al.} \cite{Bravo2007}
observed the power coefficient of a straight-bladed VAWT to become Reynolds
number independent at $Re_c \approx 4 \times 10^5$. Bachant and Wosnik
\cite{Bachant2014} observed Reynolds number independence of the power
coefficient at the optimal tip speed ratio for a high solidity cross-flow
turbine at $Re_D \approx 10^6$ or $Re_{c,\mathrm{avg}} \approx 2 \times 10^5$.

@@

Scale effects on axial-flow or horizontal-axis wind turbines are better
understood, and investigators have methods for compensating. Krogstad and
Adaramola \cite{Krogstad2012a} observed Reynolds number independence of the
performance of a 0.9 m diameter axial-flow turbine rotor at $Re_D \approx 5
\times 10^5$ in wind tunnel tests. Their turbine blades had an S826 profile
along their entire span, a profile chosen for its $Re$-independence. Walker \emph{et
al.} \cite{Walker2014} similarly chose a NACA 63-618 foil for their axial-flow
turbine tests in a towing tank, which were performed at a mid-span Reynolds
number $Re_c = 4 \times 10^5$. McTavish \emph{et al.} \cite{McTavish2013} showed how
the near-wake expansion for an axial-flow rotor is increased at higher Reynolds
numbers, concluding that physical models should be designed with
$Re$-independence in mind if they are to be run at low $Re$. It is uncertain,
however, if these methods work equally well for cross-flow turbines.

@@

\caption{\hl {University of New Hampshire (UNH)}-Reference Vertical Axis Turbine (RVAT) turbine model. Turbine blades and struts made from NACA 0020
 profiles with 0.140 m chord. Note that the upper and lower mounting flanges have
 been excluded, as these were included in the tare drag measurements.}

@@

The turbine was mounted in a frame constructed from NACA 0020 sections, shown in
Figure~\ref{fig:exp-setup}. The turbine shaft ran up through the water surface,
coupled to a Kollmorgen \hl {AKM}
%please define AKM and RPM
  permanent magnet servo \hl {motor (Manufacturer, City, Country)} with a 20:1 gearbox,
providing precise control over shaft angular velocity. This servo was controlled by the tow
tank's main motion controller for high synchronization with the carriage motion,
thereby giving precise measurement and control of the tip speed ratio. An Interface
T8 200 Nm capacity rotary torque \hl {transducer (Manufacturer, City, Country)} was installed inline between the
servo and the turbine, and the servo was also mounted on a slewing ring bearing,
which allowed a redundant measurement of torque via an arm and load cell used to
counteract the turbine moment. The frame was mounted to the carriage via linear
guides, such that the total streamwise drag force was transferred to a pair of
Sentran ZB3 500 pound-force
 capacity S-beam \hl {load cells (Manufacturer, City, Country)}, providing the rotor drag
measurements, after a separately measured tare drag was subtracted in post-processing. Similarly, a tare torque was measured by rotating the turbine shaft
in air. Turbine angular and tow carriage linear position were measured using
quadrature encoder signals, with $10^5$ counts-per-rev for the turbine and 10 ${\mu}$m resolution for the carriage position. These~signals, along with
the torque and drag signals, were sampled at 2 kHz.

Wake velocity was measured using a Nortek Vectrino+ acoustic Doppler \hl {velocimeter
(ADV)  (Manufacturer, City, Country)}, which has an approximately 6 mm diameter sampling volume and sampled at
200 Hz. The probe was mounted on an automated positioning system, also
controlled by the tow tank's main motion controller. The ADV and \hl {DAQ (please define)} systems'
sampling times were synchronized by triggering the start of data acquisition via
a pulse sent from the motion controller. Additional details of the turbine and
experimental setup are described in \cite{Bachant2015-JoT}.

@@

Approximately 1500 tows were conducted for the study reported here; each tow was
used for a single data point on either a performance curve or wake map. Each
performance curve consisted of 31~tows, where during each tow, the mean turbine
tip speed ratio was held constant, ranging from 0.1 to 3.1 in 0.1 increments. Full
performance curve data were acquired for tow speeds from 0.4 to 1.2 m/s in 0.2 m/s increments, for which the turbine diameter and approximate blade chord Reynolds
number are presented in Table~\ref{tab:Re}. Performance was also measured for
$\lambda=1.9$ at tow speeds $[0.3, 0.5, 0.7, 0.9, 1.1, 1.3]$ m/s for two tows
each.
